Generalised gravitational burst generation with Generative Adversarial Networks

### 2020-09-03 Paper_draft/main.tex
Added comments up to start of section 3

\acrodef{GW}[GW]{Gravitational wave}
\acrodef{CBC}[CBC]{compact binary coalescence} 
\acrodef{ML}[ML]{machine learning}
\acrodef{GAN}[GAN]{generative adversarial network}
\acrodef{CGAN}[CGAN]{conditional generative adversarial network}
\acrodef{ACGAN}[ACGAN]{auxilliary conditional generative adversarial network}
\acrodef{DCGAN}[DCGAN]{deep convolutional generative adversarial network}
\acrodef{CNN}[CNN]{convolutional neural networks}
\acrodef{BBH}[BBH]{binary black hole}

\section{Introduction}
%textwidth in inches: \printinunitsof{in}\prntlen{\textwidth}
\begin{comment}
\begin{itemize}
\item Need to introduce GWs - the current state of the field e.g. detections
and LVC papers \ding{51}
\item Introduce burst searches - what's the point of burst searches \ding{51} - lots of references 
\item Discuss the family of burst waveforms currently used and why - not in detail, just
an introduction \ding{51}
\item Introduce ML techniques in GWs \ding{51} - lots of references
\item What this paper does on GANs in 1 paragraph \ding{51}
\item Describe the structure of the paper 
\end{itemize}
\end{comment}

@@

% more details on burst searches
%
\ac{GW} detections so far have used a process called
matched-filtering,~\cite{Owen1998}, where a large template bank of possible
\ac{GW} waveforms are compared to the detector outputs. In order to increase
the chances of detection these template banks must span a large
multi-dimensional parameter space and the templates must be accurate, requiring
significant computational cost~\chris{the accuracy doesn't obviously lead to a
cost increase}. \ac{GW} bursts are un-modelled~\chris{except for the ones you
mentione before? SNe, glicthes, post-merger? }; therefore there are no
templates available and so matched-filtering is unsuitable for the detection of
these signals.  Instead, detection involves distinguishing the signal from
detector noise~\chris{but that's also what matched-filtering does}. Burst
searches look for excess power contained in the time-frequency
domain~\chris{all of them? In any case you have to provide references here to
the pipelines (cWB, x-pipeline, etc... Siong can help)} and rely on the
astrophysical burst waveform appearing in multiple detectors at similar times.
This is only possible if the detector noise is well characterised and the
candidate signal can be differentiated from systematic or environmental
glitches.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Generative Adversarial Networks}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{figure}
    \begin{minipage}[b]{0.3\linewidth}
        \subfloat[]{\label{fig:perceptron}\includegraphics[width=\linewidth,height=0.65\linewidth]{figures/neural_network-figure0.pdf}}
        \par
        \vspace{\baselineskip}
        \subfloat[]{\label{fig:tanh_activation}\includegraphics[width=\linewidth,height=0.65\linewidth]{figures/neural_network-figure1.pdf}}
    \end{minipage}
        \begin{minipage}[b]{0.7\linewidth}
        \subfloat[]{\label{fig:neural_network}\includegraphics[width=\linewidth,height=0.65\linewidth]{figures/neural_network-figure2.pdf}}
    \end{minipage}
    \caption{Neural Networks (a) A single neuron taking a vector of inputs and returning a singular output based on the weights, bias and activation function of the network. (b) The hyperbolic tangent used as an activation function. (c) A fully connected neural network containing two hidden layers that performs a mapping of an input vector to a singular output. \michael{Equation in the neuron looks very small}}
\end{figure}

Since the birth of programmable computers, people have wondered if machines can develop true intelligence beyond formal mathematical procedures. Today, ML aims to learn apparent relationships held within the given data or `training data' in order to make accurate predictions without the need for additional programming . A common approach in ML relies on the computer learning on past experience to make decisions on future events. \michael{Not sure about the use of computer here, maybe
\textit{relies on a model}?} The hierarchical nature of this approach means that the machine can build complicated concepts from simpler ones. This branch of AI is called deep learning.
Neural networks are the quintessential machine learning algorithm that aims to approximate a function. They are built from many single processing units called neurons. The simpliest Neural Network is the perceptron layer \cref{fig:perceptron} which holds a single neuron that takes several real inputs $x_{1}, ... , x_{i}$ and maps them to an output according to the following linear equation:

\begin{equation}
\sigma(\sum_i w_i x_i + b),
\label{eqn:neuron}
\end{equation}
where $w$ and $b$ are the weights and bias and $\sigma$ denotes the activation function. The weights are numbers which can be thought as the strength between connections. The output of a neuron is called its activation. Different choices of activation functions allow the user to simulate various models. A common example of an activation function is the hyperbolic tangent \cref{fig:tanh_activation} that is used when the outputs of the network must be both positive and negative and re-scaled to $[-1,1]$. It is often useful to introduce a bias, $b$, such that the neuron remains inactive above zero but is active when the sum reaches a defined threshold. This bias is added before the activation function and it tells us how high the weighted sum needs to be before the neuron becomes active.

The output of a single neuron is defined by \cref{eqn:neuron} and gives a prediction $\hat{y}$ that can be compared to the real value $y$ through a cost or loss function. If the loss is non-zero, the network must work to minimise this function by updating the weights in the negative direction of the loss gradient in a process referred to as gradient descent. This loss function plays a critical role in how neural networks learn.

A neural network contains many single neurons connected in a layered structure as shown in \cref{fig:neural_network}. The activations of the first layer (or input layer) act as the inputs to the second layer and so on until the output layer. Multi-layered neural networks have intermediate layers between the input and output stages dubbed the hidden layers as the computations performed are not accessible to the user. The network can be thought of as a function F: $\mathbb{R}^N\rightarrow \mathbb{R}^M $ reliant on how activations from one layer bring activations in the next. The system is analogous to biology, where, some groups of neurons in animal brains cause certain others to fire.

\subsection{Convolutional Neural Networks}
Convolutional Neural Networks (CNNs) are designed to work with grid-like structures that exhibit strong local spatial dependencies. An example would be a two-dimensional image as each coloured pixel has similar intensities to their near neighbours. The value contained in each pixel can be thought of as adding an additional dimension or channel to the image. Although most work with CNNs involve image-based data, they can be applied to other spatially adjacent data types such as time-series and
text items. CNNs are defined by the use of a convolution operation, a mathematical operation that expresses the amount overlap between functions. The convolution is applied by shifting one grid-like structure over the other, drawing out spatially important features between the two.  A strided convolution reduces the dimensionality of the input while retaining spatial features \michael{How? This is core part of the architecture of the GAN but is quite vague, I don't think it's
obvious that stride imples moving the filter further. Maybe something like: \textit{the distance by which the grid is shifted is known as the stride and increasing it reduces the dimensionality of the output}}. This can result in a loss of information along the boarders. To combat this, the input is padded with zeros at the edges which does not interfere with the convolutional operation. The output of the convolutional layer is then passed to an activation function.

\subsection{GANs}

A subset of deep learning that has seen fruitful development in recent years is \acp{GAN} \cite{Goodfellow2014}. These unsupervised algorithms learn
patterns in a given training data set using an adversarial process. The
generations from \acp{GAN} are state-of-the-art in fields such as high quality image
fidelity \cite{brock2018large,karras2019analyzing}, text-to-image translation
\cite{reed2016generative} and video prediction \cite{liang2017dual} as well as
time series generations \cite{esteban2017realvalued}.

% How a GAN works
%
\acp{GAN} train two competing neural networks, consisting of a discriminator
that is set up to distinguish between real and fake data and a generator that
produces synthetic reproductions of the real data. \michael{Is it clear to the reader that the \textit{fake} data is the same as the synthetic data?} The generator performs a
mapping from an input noise vector $\mathbf{z}$, that is usually sampled from a n-dimensional Gaussian space known as the latent space, to its representation of the
data and the discriminator maps its input $\mathbf{x}$ to a probability that
the input came form either the training data or generator.  During training,
the discriminator is given a batch of samples that contains one half real data
and one half fake data which it then makes predictions on. The loss for the
discriminator is calculated by comparing its predictions to the labelled data
through the binary cross-entropy function. \jordan{note to myself: This is important since eqn 2 is derived from it (thats where the logs come from)} The training process of a \ac{GAN}
alternatively updates the weights of the discriminator and generator based on
information on its competitors loss function. This loss of discriminator is
used to update the weights of the generator to produce more realistic samples
of the input distribution, the loss of the generator encourages the
discriminator to update its classification abilities. GANs seek to find an equilibrium between the generator and discriminator in the form of a two-player game that can be summarised by the following:

\begin{comment}
Both networks compete in
a minimax game~\chris{what is a minimax game? references?} which the generator
is trying to minimise and the discriminator is trying to
maximise:~\chris{nicely written, however, it needs an extra 25\% explanation
for people not familar with ML or GANs in particular. Try to spell things out
more. Also refer to Fig 1 somewhere in this section.}
\end{comment}
%
\begin{equation}
  \mathop{\text{min}}_{G}  \mathop{\text{max}}_{D} V(D,G) = \mathbb{E}_{\mathbf{x} \sim p_{\text{r}}(\mathbf{x})} [\text{log} D(\mathbf{x})] \\ + \mathbb{E}_{\mathbf{z} \sim p_{\text{z}}(\mathbf{z})} [\text{log}(1-D(G(\mathbf{z})))],
\label{equation:GANloss}
\end{equation}

where V is an objective function that is to be optimized. Objective functions are a general term for either a loss function (to be minimized) or a negative loss function (to be maximised). D is the discriminator, G is the generator and ${\mathbf{x} \sim p_{\text{r}}(\mathbf{x})}$, ${\mathbf{z} \sim p_{\text{z}}(\mathbf{z})}$ are samples taken from the training data and generated data respectfully. In practice this approach was found to be unstable \michael{Cite?}. Early on in training the
generators weights are initialised randomly meaning that the early generations are noisy and uninformative. This means that the discriminator can easily learn to spot the fake generations and so the discriminator wins and the generator cannot continue to improve. To overcome this the framing of the generator is often changed. Rather than minimizing $\text{log}(1-D(G(\mathbf{z})))$ the objective now is to maximise $\text{log}(D(G(\mathbf{z})))$, that is, to maximise the probability of the
generations being predicted as real. The idea being that we always want the losing side to be able to leverage \michael{information?} from its competitor. 
%

% describing the problems with GANs 
%
In theory, the adversarial process will eventually lead to the local Nash
equilibrium \cite{Nash1950} whereby both neural networks are trained
optimally. In practice, however, \acp{GAN} are notoriously difficult
to train. Such difficulties include: Non-convergence, where the model
parameters oscillate and the loss never converges, mode collapse where the generator produces a
limited diversity of samples, and diminishing
gradients when applying gradient descent to a loss function that is discontinuous.

\subsection{Conditional GANs}

% introduce CGANs and ACGANs
%
To gain more control over what a GAN is able to generate, a conditional variant of \acp{GAN} named \acp{CGAN}~\cite{cgan} was introduced by feeding in extra information into the generator and discriminator such as a class label or attribute label, c. The objective function \cref{equation:GANloss}  is then modified:
\begin{equation}
  \mathop{\text{min}}_{G}  \mathop{\text{max}}_{D} V(D,G) = \mathbb{E}_{\mathbf{x,c} \sim p_{\text{r}}(\mathbf{x,c})} [\text{log} D(\mathbf{x,c})] \\ + \mathbb{E}_{\mathbf{z} \sim p_{\text{z}}(\mathbf{z}),\mathbf{c} \sim p_{\text{r}}(\mathbf{c})} [\text{log}(1-D(G(\mathbf{z,c})))].
\label{equation:cGANloss}
\end{equation}

This simple addition has shown to work well in practice, for instance in image-to-image translation \cite{isola2016imagetoimage}. We will be using a conditional GAN for this study.
%
\begin{comment}adds structure to the latent space by providing the
generator with a class or attribute label. The generator learns to segment the
latent space by clustering distributions which have similar properties, making
a point in latent space conditional on a class~\chris{are we sure about this?
all classes share the same latent space and I don't think that the space gets
segmented into areas corresponding to different classes - I do think that it
gets broken down into spaces with different intrinsic physical properties e.g.,
high frequency, long duration, etc...}.
\end{comment}
%
\begin{comment}
This idea was extended further with
\acp{ACGAN}~\cite{odena2016conditional} that require the discriminator to
output a probability of data belonging to each class. A pictorial
representation on the differences between these approaches is shown in
Fig.~\ref{fig:gan_comparison} \jordan{self note: Probably move away from ACGAN part, standalone aux classifiers can be trained later if that's what you want}. 
\end{comment}

\begin{figure}
    \centering
        \input{figures/GAN_diagram_vertical}%     without .tex extension
        \input{figures/ACGAN_diagram_vertical}%     without .tex extension
    \caption{Comparison of the original GAN method and the
Conditional-GAN method. For CGANs the training data requires a label denoting
its class that is also fed to the generator which then learns to generate
waveforms based on the input label.} \label{fig:gan_comparison}
\end{figure}

>>>>

\acrodef{GW}[GW]{gravitational wave}
\acrodef{CBC}[CBC]{compact binary coalescence} 
\acrodef{ML}[ML]{machine learning}
\acrodef{AI}[AI]{artificial intelligence}
\acrodef{CNN}[CNN]{convolutional neural network}
\acrodef{GAN}[GAN]{generative adversarial network}
\acrodef{CGAN}[CGAN]{conditional generative adversarial network}
\acrodef{ACGAN}[ACGAN]{auxilliary conditional generative adversarial network}
\acrodef{DCGAN}[DCGAN]{deep convolutional generative adversarial network}
\acrodef{CNN}[CNN]{convolutional neural networks}
\acrodef{BBH}[BBH]{binary black hole}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%textwidth in inches: \printinunitsof{in}\prntlen{\textwidth}
\begin{comment}
\begin{itemize}
\item Need to introduce GWs - the current state of the field e.g. detections
and LVC papers \ding{51}
\item Introduce burst searches - what's the point of burst searches \ding{51} - lots of references 
\item Discuss the family of burst waveforms currently used and why - not in detail, just
an introduction \ding{51}
\item Introduce ML techniques in GWs \ding{51} - lots of references
\item What this paper does on GANs in 1 paragraph \ding{51}
\item Describe the structure of the paper 
\end{itemize}
\end{comment}

@@

% more details on burst searches
%
\ac{GW} searches for modelled signals so far have used a process called
matched-filtering,~\cite{Owen1998}, where a large template bank of possible
\ac{GW} waveforms are compared to the detector outputs. In order to increase
the chances of detection these template banks must span a large
multi-dimensional parameter space and the templates must be accurate, requiring
significant computational cost~\chris{the accuracy doesn't obviously lead to a
cost increase}. \ac{GW} bursts are un-modelled~\chris{except for the ones you
mentione before? SNe, glicthes, post-merger? }; therefore there are no
templates available and so matched-filtering is unsuitable for the detection of
these signals.  Instead, detection involves distinguishing the signal from
detector noise~\chris{but that's also what matched-filtering does}. Burst
searches look for excess power contained in the time-frequency
domain~\chris{all of them? In any case you have to provide references here to
the pipelines (cWB, x-pipeline, etc... Siong can help)} and rely on the
astrophysical burst waveform appearing in multiple detectors at similar times.
This is only possible if the detector noise is well characterised and the
candidate signal can be differentiated from systematic or environmental
glitches.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Generative Adversarial Networks}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{figure}
    \begin{minipage}[b]{0.3\linewidth}
        \subfloat[]{\label{fig:perceptron}\includegraphics[width=\linewidth,height=0.65\linewidth]{figures/neural_network-figure0.pdf}}
        \par
        \vspace{\baselineskip}
        \subfloat[]{\label{fig:tanh_activation}\includegraphics[width=\linewidth,height=0.65\linewidth]{figures/neural_network-figure1.pdf}}
    \end{minipage}
        \begin{minipage}[b]{0.7\linewidth}
        \subfloat[]{\label{fig:neural_network}\includegraphics[width=\linewidth,height=0.65\linewidth]{figures/neural_network-figure2.pdf}}
    \end{minipage}
    \caption{Neural Networks (a) A single neuron taking a vector of inputs and
returning a single output based on the weights, bias and activation function
of the network. (b) The hyperbolic tangent used as an activation
function.~\chris{is this the actual activation that you use in the GAN?
otherwise why focus on it alone, you could plot a handful on the same plot} (c)
A fully connected~\chris{you haven't mentioned what a fully connected network
actually is in the text} neural network containing two hidden layers that performs a
mapping of an input vector to a single output. \michael{Equation in the
neuron looks very small}~\chris{I agree, but the equation in (b) is even
smaller. I would also recommed that you make the style of the FC network hidden
and output layer neurons the same as the single perceptron in terms of colours.
You could put $x_i$s in the input layer.}}
\end{figure}

% inroduce the concept of ML
%
Since the birth of programmable computers, people have wondered if machines can
develop true intelligence beyond formal mathematical procedures. Today, \ac{ML}
aims to learn apparent relationships held within the given data or `training
data' in order to make accurate predictions without the need for additional
programming. A common approach in \ac{ML} relies on the computer learning on
past experience to make decisions on future events. \michael{Not sure about the
use of computer here, maybe \textit{relies on a model}?} The hierarchical
nature of this approach means that the machine can build complicated concepts
from simpler ones~\chris{what hierarchical nature? You haven't mentioned
anything about that.}. This branch of \ac{AI} is called deep
learning~\chris{OK, but as long as you explain the hierarchy thing and how it
implies complicated concepts from simpler ones.}.

% Introduce basic neural networks - a perceptron layer
%
Neural networks are the quintessential~\chris{really? quintessential?} \ac{ML}
algorithm that aims to approximate a function. They are built from many single
processing units called neurons. The simpliest Neural Network is the perceptron
layer~\cref{fig:perceptron} which holds a single neuron that takes several real
inputs $x_{1},\ldots, x_{i}$~\chris{and outputs what? also $i$ is the index and not the "final" index}
and maps them to an output according to the linear equation 
%
\begin{align}
\sigma(\sum_i w_i x_i + b),
\label{eqn:neuron}
\end{align}
%
\chris{it's not an equation if it's not equal to something. Please make it
eaqual something.} where $w$ and $b$ are the weights and bias and $\sigma$
denotes the activation function. The weights are numbers which can be thought
as the strength between connections~\chris{what connestions? between what
objects?}. The output of a neuron is called its activation and different
choices of activation function allow the user~\chris{are we users or developers
or what?} to simulate various models. A common example of an activation
function is the hyperbolic tangent~\cref{fig:tanh_activation} that is used when
the outputs of the network must be both positive and negative~\chris{how can an
output be both positive and negative. The language is ambiguous and confusing}
and re-scaled to $[-1,1]$~\chris{at this stage in the explanation the reader
has no idea what an output would look like or why it would be rescaled}. It is
often useful to introduce a bias, $b$, such that the neuron remains inactive
above zero but is active when the sum reaches a defined threshold. This bias is
added before the activation function and it tells us how high the weighted sum
needs to be before the neuron becomes active.~\chris{well kind of, but also not
really. I would advise you to rewrite this (if we keep it) in a minimalist
style just explaining what each item is and not trying to describe what it
does.}

% the cost or loss function
%
The output of a single neuron is defined by~\cref{eqn:neuron} and gives a
prediction $\hat{y}$ that can be compared to the real value $y$~\chris{what the
hell is $y$, it's not been mentioned before} through a loss (also known as a
cost) function. If the loss is non-zero~\chris{no, it doesn't matter whether
the number is above or below or equal to zero, the machine will try to minimise
the number (if negative, it will try to make it more negative)}, the network
must work to minimise this function by updating the weights in the negative
direction of the loss gradient in a process referred to as gradient
descent~\chris{give reference and also make the distinction between this and
the actual method used to do it, back progagation}. This loss function plays a
critical role in how neural networks learn~\chris{OK, but are you going to
expand on that or just leave it there? As far as I can tell from the text it's
the *only* way network's learn so that's pretty critical and already evident}.

% basic network structure
%
A neural network contains many single neurons connected in a layered structure
as shown in~\cref{fig:neural_network}. The activations of the first layer (or
input layer) act as the inputs to the second layer and so on until the output
layer. Multi-layered neural networks have intermediate layers between the input
and output stages dubbed the hidden layers as the computations performed are
not accessible to the user~\chris{well that's not really true. We can access
everything if we want but it depends on what you define as a user. Best to
simply state what they're clalled and no need to give a reason.}. The network
can be thought of as a function F: $\mathbb{R}^N\rightarrow
\mathbb{R}^M$~\chris{does F not have to be in math-mode too? Is is a
mathematical symbol?} reliant on how activations from one layer bring
activations in the next~\chris{yes, kind of, but also reliant on the trained
weights and biases too, so actually reliant on everything}. The system is
analogous to biology, where, some groups of neurons in animal brains cause
certain others to fire~\chris{best to stay away from analogies like these where
the practical specifics can be picked apart}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Convolutional Neural Networks}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\acp{CNN} are designed to work with grid-like input structures that exhibit
strong local spatial dependencies. An example would be a two-dimensional image
as each coloured pixel has similar intensities to their near
neighbours~\chris{it's not about that really, it's like you said, it's about
spatial relationships}. The value contained in each pixel can be thought of as
adding an additional dimension or channel to the image~\chris{no, the dimension
and channel are specifically defined. A 2D image has 2 dimensions. An RGB image
has 3 numbers per pixel and these are typcally stored as channels. Channels are
dealt with differently to dimensions}. Although most work with \acp{CNN}
involve image-based data, they can be applied to other spatially adjacent data
types such as time-series and text items. \acp{CNN} are defined by the use of a
convolution operation, a mathematical operation that expresses the amount
overlap between functions. The convolution is applied by shifting one grid-like
structure over the other, drawing out spatially important features between the
two~\chris{yes, but what is being dragged over what? Time to define the
convolutional neuron? Maybe add an equation?}. A strided convolution reduces
the dimensionality of the input while retaining spatial features \michael{How?
This is core part of the architecture of the GAN but is quite vague, I don't
think it's obvious that stride imples moving the filter further. Maybe
something like: \textit{the distance by which the grid is shifted is known as
the stride and increasing it reduces the dimensionality of the
output}}~\chris{why are we talking about strides at this point? Are they
important for the GAN? You haven't mentioned kernel sizes or dropout. are these
equally important?}. This can result in a loss of information along the
borders~\chris{the borders of what?}. To combat this, the input is padded with
zeros at the edges which does not interfere with the convolutional operation.
The output of the convolutional layer is then passed to an activation function.

~\chris{I'm getting the sense that these 2 sections of ANNs and CNNs aren't
really helpful in their current state. They should either be removed OR tidied
up. I don't think they need to be expanded upon but various statements can be
removed and other important things can be added. Imagine reading this with no
knowledge of ML. Do you get informed by this or confused?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{GANs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% basic intro to GANs
%
A subset of deep learning that has seen fruitful development in recent years
are \acp{GAN}~\cite{Goodfellow2014}. These unsupervised~\chris{they have
training data don't they? Is that unsupervised?} algorithms learn patterns in a
given training data set using an adversarial process. The generations from
\acp{GAN} are state-of-the-art~\chris{a phrase that will not stand the test of
time} in fields such as high quality image
fidelity~\cite{brock2018large,karras2019analyzing}, text-to-image
translation~\cite{reed2016generative}, and video
prediction~\cite{liang2017dual} as well as time series
generations~\cite{esteban2017realvalued}~\chris{do we still not have any other
GW GAN applications to reference or are we hoping to still be the first?}.

% Basic components of a GAN
%
\acp{GAN} train two competing neural networks, consisting of a discriminator
network that is set up to distinguish between real and fake data and a
generator network that produces synthetic (fake) versions of the real data.
\michael{Is it clear to the reader that the \textit{fake} data is the same as
the synthetic data?} The generator performs a mapping from an input noise
vector $\mathbf{z}$, that is usually sampled from a $n$-dimensional~\chris{be
careful with statments about $n$-dimensions since $n$ is a variable and could
mean the number of latent dimensions or the number of class dimensions or data
dimensions, etc...} Gaussian space known as the latent space, to its
representation of the data, and the discriminator maps its input $\mathbf{x}$ to
a probability that the input came from either the training (real) data or
generator (fake).

% training a GAN
%
During training, the discriminator is given a batch~\chris{you have not
introduced batches yet. Best to do that in the ANN section} of samples that
contains one half real data and one half fake data which it then makes
predictions on.  The loss for the discriminator is calculated by comparing its
predictions to the labelled data through the binary cross-entropy function.
\jordan{note to myself: This is important since eqn 2 is derived from it (thats
where the logs come from)}~\chris{either give a reference to binary cross
entropy or define it here mathematically}. The training process of a \ac{GAN}
alternatively updates the weights of the discriminator and generator based on
information on its competitors loss function~\chris{you haven't fully explained
how the loss is computed in the 2 phases of each training pass. I recommend
that you add that here other wise discussion of generator loss and discrimator
loss makes no sense.}. This loss of the discriminator is used to update the
weights of the generator to produce more realistic samples of the input
distribution, the loss of the generator encourages the discriminator to update
its classification abilities. \acp{GAN} seek to find an equilibrium between the
generator and discriminator in the form of a two-player game that can be
summarised by 
%
\begin{align}
  \mathop{\text{min}}_{G}  \mathop{\text{max}}_{D} V(D,G) &= \mathbb{E}_{\mathbf{x} \sim p_{\text{r}}(\mathbf{x})} [\text{log} D(\mathbf{x})] + \mathbb{E}_{\mathbf{z} \sim p_{\text{z}}(\mathbf{z})} [\text{log}(1-D(G(\mathbf{z})))],
\label{equation:GANloss}
\end{align}
%
% remove these comments 
%
Both networks compete in
a minimax game~\chris{what is a minimax game? references?} which the generator
is trying to minimise and the discriminator is trying to
maximise:~\chris{nicely written, however, it needs an extra 25\% explanation
for people not familar with ML or GANs in particular. Try to spell things out
more. Also refer to Fig 1 somewhere in this section.}
\end{comment}
%
where $V$ is an objective function that is to be optimised. Objective functions
are a general term for either a loss function (to be minimized) or a negative
loss function (to be maximised). We use $D$ to define the discriminator
function~\chris{output interpreted as a probability?} , $G$ for the generator
function and ${\mathbf{x} \sim p_{\text{r}}(\mathbf{x})}$, ${\mathbf{z} \sim
p_{\text{z}}(\mathbf{z})}$ are samples taken from the training data and
generated data respectfully~\chris{no, $z$ can't be from the data since it's
not called $x$. It is from the latent space distribution.}. In practice this
approach was found to be unstable \michael{Cite?}~\chris{unstable for us or
unstable in general, also add reference.}. Early on in training the generators
weights are initialised randomly meaning that the early generations are noisy
and uninformative. This means that the discriminator can easily learn to spot
the fake generations and so the discriminator wins and the generator cannot
continue to improve. To overcome this the framing~\chris{what is framing in
this context?} of the generator is often changed. Rather than minimizing
$\text{log}(1-D(G(\mathbf{z})))$ the objective now is to maximise
$\text{log}(D(G(\mathbf{z})))$, that is, to maximise the probability of the
generations being predicted as real. The idea being that we always want the
losing side to be able to leverage \michael{information?} from its
competitor.~\chris{OK, does this all mean that the previous equation is wrong?
If so, why quote it. Just use the actual correct equation and explain why it's
correct rather than explaining why something is wrong. I would also try to link
up this equation with how we actually train i.e., we compute the first term by
drawing samples from the training set $x$ and getting the average log-prob of
the discriminator (not quite true since we actually input a mix of real $x$ and
fake $G(z)\rightarrow x'$). Then we take a set of fake $x'$ made using $G(z)$
with $z$ drawn from the latent space, and get the loss (but we switch the
labels hence the $1-D(G(z))$. See what I mean? You shouls also refer to Fig 2
here.} 
%

% describing the problems with GANs 
%
In theory, the adversarial process will eventually lead to the local Nash
equilibrium~\cite{Nash1950} whereby both neural networks are trained
optimally~\chris{vague and not informative statement. What does optimal mean
here?}. In practice, however, \acp{GAN} are notoriously difficult to train with
difficulties that include: Non-convergence, where the model parameters
oscillate and the loss never converges, mode collapse where the generator
produces a limited diversity of samples, and diminishing gradients when
applying gradient descent to a loss function that is discontinuous.~\chris{OK,
difficult to train. I understand. Why are you telling the reader this? Will we
be referring back to these issues later?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Conditional GANs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% introduce CGANs and ACGANs
%
To gain more control over what a GAN is able to generate, a conditional variant
of \acp{GAN} named \acp{CGAN}~\cite{cgan} was introduced by feeding in extra
information into the generator and discriminator such as a class label or
attribute label, $c$. The objective function~\cref{equation:GANloss}~\chris{I
think it owuld be be best if you just stick to loss function rather than
switching between loss and objective} is then
%
\begin{align}
  \mathop{\text{min}}_{G}  \mathop{\text{max}}_{D} V(D,G) &= \mathbb{E}_{\mathbf{x,c} \sim p_{\text{r}}(\mathbf{x,c})} [\text{log} D(\mathbf{x,c})] + \mathbb{E}_{\mathbf{z} \sim p_{\text{z}}(\mathbf{z}),\mathbf{c} \sim p_{\text{r}}(\mathbf{c})} [\text{log}(1-D(G(\mathbf{z,c})))].
\label{equation:cGANloss}
\end{align}

This simple addition has shown to work well in practice, for instance in
image-to-image translation~\cite{isola2016imagetoimage}. We will be using a
conditional GAN for this study.~\chris{OK, but you have not defined the new
changes to the loss function in the text. Please indicate that the training
data and labels are being drawn from a joint distribution $p_r(x,c)$ and when
making fake data you are drawing from $p_r(c)$ independent of =$p(z)$. However,
what does the $r$ stand for and why is it important, and what defines $p_r(c)$?}

~\chris{Finally, you need to refer to and discuss Figure 2. It is currently
neither referenced or discussed.}

% get rid of these comments if they are no longer useful
%
\begin{comment}adds structure to the latent space by providing the
generator with a class or attribute label. The generator learns to segment the
latent space by clustering distributions which have similar properties, making
a point in latent space conditional on a class~\chris{are we sure about this?
all classes share the same latent space and I don't think that the space gets
segmented into areas corresponding to different classes - I do think that it
gets broken down into spaces with different intrinsic physical properties e.g.,
high frequency, long duration, etc...}.
\end{comment}
%
\begin{comment}
This idea was extended further with
\acp{ACGAN}~\cite{odena2016conditional} that require the discriminator to
output a probability of data belonging to each class. A pictorial
representation on the differences between these approaches is shown in
Fig.~\ref{fig:gan_comparison} \jordan{self note: Probably move away from ACGAN part, standalone aux classifiers can be trained later if that's what you want}. 
\end{comment}

\begin{figure}
    \centering
        \input{figures/GAN_diagram_vertical}%     without .tex extension
        \input{figures/ACGAN_diagram_vertical}%     without .tex extension
    \caption{Comparison of the original GAN method and the
Conditional-GAN method. For CGANs the training data requires a label denoting
its class that is also fed to the generator which then learns to generate
waveforms based on the input label.~\chris{OK, a few things. Try to separate
these a bit and maybe make them subplots with sublabels indicating GAN and CGAN
distinction. Replace noise with latent (or otherwise). Noise implies GW data
noise which it is not. Overall it's a bit basic and doesn't indicate either the
2 competing training steps or any aspect of the loss or specify that $x$ and
$z$ (and $c$) are drawn from distributions already defined in the text. Maybe it's OK without
that. As for the CGAN, I'm troubled by the fact that there is only one $c$ box.
It implies that the same $c$ value is given to each real and fake $x$ data when
in fact each $x$ sample has it's own randomly drawn $c$ value.}} \label{fig:gan_comparison}
\end{figure}
